---
title: "Cereals Consumption by Cereal Type"
author: "David Jackson"
date: "3/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(scales)
library(dplyr)
library(tidyr)
library(readr)
library(skimr)
```

## Cereals Consumption by Cereal Type
"Breakdown of the average per capita intake of cereals by specific cereal-based commodity types,
measured in kilocalories per person per day. This figure measures the primary equivalent of all food
products derived from a given commodity (e.g. "wheat" represents the primary equivalent of its derived
products). Data refers to cereal food supply at the consumer level but does not account for consumer
wastage."

* Source: https://ourworldindata.org/grapher/per-capita-consumption-of-cereals-by-commodity-type-daily-kilocalories

```{r}
cereals <- read.csv("./DATA/per-capita-consumption-of-cereals-by-commodity-type-daily-kilocalories.csv") 
colnames(cereals) <- c("Country","Code","Year","Wheat","Rice","Barley","Maize","Rye","Oats","Sorghum")
summary(cereals)
```

#### Pull Out World Numbers

```{r}
world_longer <- cereals %>% filter(Country =="World") %>% select(-Code) %>%
  pivot_longer(cols = Wheat:Sorghum)
```

#### Plot of Worldwide 

```{r}
world_yearly <- world_longer %>% group_by(Year) %>%
  summarise(Total = sum(value),
            Median = median(value),
            Mean = mean(value))
```
```{r}
ggplot(world_yearly) + geom_line(aes(x=Year,y=Total)) +
  labs(title="Cereals: kilocalories per person per day",y="kcal/capita/day") + 
    scale_y_continuous(labels = comma)
```
```{r}
ggplot(world_yearly) + geom_line(aes(x=Year,y=Mean)) +
  labs(title="Cereals: Mean kilocalories per person per day",y=" Mean kcal/capita/day") + 
    scale_y_continuous(labels = comma)
```
