---
title: "Cereals Consumption by Cereal Type"
author: "David Jackson"
date: "3/09/2021 9:20 PM/EST"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(scales)
library(dplyr)
library(tidyr)
library(readr)
library(skimr)
```

## Cereals Consumption by Cereal Type
"Breakdown of the average per capita intake of cereals by specific cereal-based commodity types,measured in kilocalories per person per day. This figure measures the primary equivalent of all food products derived from a given commodity (e.g. "wheat" represents the primary equivalent of its derived
products). Data refers to cereal food supply at the consumer level but does not account for consumer wastage."

* Source: https://ourworldindata.org/grapher/per-capita-consumption-of-cereals-by-commodity-type-daily-kilocalories

```{r}
cereals <- read.csv("./DATA/per-capita-consumption-of-cereals-by-commodity-type-daily-kilocalories.csv") 
colnames(cereals) <- c("Country","Code","Year","Wheat","Rice","Barley","Maize","Rye","Oats","Sorghum")
summary(cereals)
```

#### Pull Out World Numbers

```{r}
world_longer <- cereals %>% filter(Country =="World") %>% select(-Code) %>%
  pivot_longer(cols = Wheat:Sorghum)
```

#### Plot of Worldwide 

```{r}
world_yearly <- world_longer %>% group_by(Year) %>%
  summarise(Total = sum(value),
            Median = median(value),
            Mean = mean(value))
```
```{r}
ggplot(world_yearly) + geom_line(aes(x=Year,y=Total),lwd=1.5,col='red') +
  labs(title="Worldwide Cereals: kilocalories per person per day",y=" (x1000)Calories/capita/day") + 
    scale_y_continuous(labels = comma)
```
```{r}
ggplot(world_yearly) + geom_line(aes(x=Year,y=Mean),lwd=1.5,col="darkblue") +
  labs(title="Worldwide Cereals: Mean kilocalories per person per day",y=" Mean Calories(x1000)/capita/day") + 
    scale_y_continuous(labels = comma)
```

### KiloCalories by Cereal Type and Year

```{r}
cereals_longer <- cereals %>% filter(Country !="World") %>% select(-Code) %>%
  pivot_longer(cols = Wheat:Sorghum) 

```
```{r}
 cereals_by_years <- cereals_longer %>% filter(value >0) %>% group_by(name,Year) %>% 
  summarise(Count = n(),
            Total = sum(value),
            Mean = mean(value)) %>% ungroup()
```
```{r}
ggplot(cereals_by_years) + geom_line(aes(x=Year,y=Total,col=name),lwd=1) +
  scale_y_log10(labels = comma) + 
  labs(title="Total kilocalories Per Year by Cereal Type")
```

```{r}
ggplot(cereals_by_years) + geom_line(aes(x=Year,y=Total),lwd=1) + 
  facet_wrap(~name,scale = "free_y",ncol=2) +
  scale_y_log10(labels = comma) + 
  labs(title="Total kilocalories Consumed Per Year by Cereal Type")
```
